Integrating brain and biomechanical models - a new paradigm for understanding neuro-muscular control

### 2017-09-01 papers/integration/omm_integ.tex
Updated, ready to submit paper

\section{}
To date, realistic models of how the central nervous system governs
behaviour have been restricted in scope to the brain, brainstem or
spinal column, as if these existed as disembodied organs.
Further, the model is often exercised in relation to an in vivo
physiological experiment with input comprising an impulse, a periodic
signal or constant activation, and output as a pattern of neural
activity in one or more neural populations. Any link to behaviour is
inferred only indirectly via these activity patterns.
%
We argue that to discover the principles of operation of neural
systems, it is necessary to express their behaviour in terms of
physical movements of a realistic motor system, and to supply inputs
that mimic sensory experience. To do this with confidence, we must
connect our brain models to neuro-muscular models and provide relevant
visual and proprioceptive feedback signals, thereby closing the loop
of the simulation.
%
This paper describes an effort to develop just such an integrated
brain and biomechanical system using a number of pre-existing
models. It describes a model of the saccadic oculomotor system
incorporating a neuromuscular model of the eye and its six extraocular
muscles. The position of the eye determines how illumination of a
retinotopic input population projects information about the location
of a saccade target into the system.
%
A pre-existing saccadic burst generator model was incorporated into
the system, which generated motoneuron activity patterns suitable for
driving the biomechanical eye. The model was demonstrated to make
accurate saccades to a target luminance under a set of environmental
constraints.
%
Challenges encountered in the development of this model showed the
importance of this integrated modelling approach. Thus, we exposed
shortcomings in individual model components which were only apparent
when these were supplied with the more plausible inputs available in a
closed loop design. Consequently we were able to suggest missing
functionality which the system would require to reproduce more
realistic behaviour.
%
The construction of such closed-loop animal models constitutes a new
paradigm of {\em computational neurobehaviour} and promises a more
thoroughgoing approach to our understanding of the brain's function as
a controller for movement and behaviour.

@@

% New introduction, tailored to the paradigm shift title.
(New Introduction:)

The field of computational neuroscience has provided
many \e{systems models} of the brain~\cite{arai_two-dimensional_1994,gancarz_neural_1998,hazy_towards_2007,blenkinsop_frequency_2017}. We
refer to these as \e{mechanistic computational models}, meaning
models which consist of populations of neural elements, interconnected
in a biologically plausible manner, which simulate the operation of
the brain.

Whilst they differ in scale and complexity, these models all seek to
describe the fundamental mechanisms behind common animal behaviours
such as locomotion, threat evasion, reaching or feeding. However,
almost none of the models cited here actually reproduce these
behaviours. In each case, the activity in a certain population of
neurons is taken to be representative of a behavioural outcome.

\cmnt{Background, gap, solution (cheers Sean!)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
(Old introduction, which became the VPH mini-paper intro:)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Can make comparisons
As mechanistic computational models of brain systems become more
sophisticated we can aspire to make realistic comparisons between the
behaviour of the model and that of the animal or human.

% Short desc of mechanistic computational model
By \e{mechanistic computational model}, we mean a model which
consists of populations of neural elements, interconnected in a
biologically plausible manner. Such a model may consist of neural
elements that produce a time-dependent membrane voltage and realistic
spiking behaviour (Hodgkin-Huxley, Fizhugh-Nagumo, Izhikevich, Leaky
integrate and fire) in which case the neural element's activity can be
obtained by counting action potentials (or considering their
timing). Alternatively it may (as it does in this work) consist of
rate-coded neural elements, whose activation variable is related to
the spiking rate of a number of neurons (refs).

% Sometimes, ok to just have a brain model
In some cases, it is sufficient to take the activity of an internal
population within the brain model as being representative of the
induced behaviour. For example, a choice made in a \e{go/no-go} task
could be determined from activity in a population within a Basal
Ganglia model
(\cite{nambu_discharge_1990,kuhn_event-related_2004}). The decision
to \e{go} is selected by a reduction of activity in this population;
maintenance of activity implies \e{no-go}. Error rates in this task
could be compared with experimentally determined error rates in
primate subjects.
% Brain model comparison validates model
If the expression of behaviour by the model is similar when compared
with the experimental data, this serves to validate the model for the
healthy, neuro-typical subject.
% Now describe comparison between model and expt. for diseased
% subject.
If we have confidence in the model's description of the behaviour for
the healthy subject, we can now progress to a comparison between the
model and an unhealthy subject. If we can identify a characteristic
degradation in an aspect of the brain for an unhealthy subject, then
we can seek to modify the model to mirror this degradation. By
comparing the relationship between the model's behaviour as a function
of the degradation parameter with the unhealthy individual's behaviour
as a function of disease progression, we may be able to relate the
disease progression state to the brain parameter and predict the
time-course of the internal degradation of the brain.
% Leave further discussion along these lines to the discussion
% section.

% But sometimes you might need brain and neuromuscular model
Many behaviours are more complex and subtle than a decision. In the
disease state, it may be the dynamics of the movement which is
symptomatic rather than its outcome. A diagnostic task may produce an
outcome whose error rate varies with disease state, but it may also
produce movement characteristics which correlate with the disease
progression such as the slow, jerky or inaccurate movements seen in
Parkinson's Disease. In order to determine the validity of a model
which seeks to describe more complex movements, it is necessary to
express the movement commanded by the brain network as a realistic,
simulated motion by means of a biomechanical model, driven by the
brain model.

% Motivation here is to study movement disorders by means of a
% comparison between healthy subjects, the healthy model and diseased
% subjects and the diseased model.
Our motivation is to study movement disorders using our brain
models. Because the challenge of integrating brain and biomechanical
models has not, to our knowledge, previously been addressed, we chose
to express some of the simplest, and most studied movements made by
vertebrates: the rapid eye movements known as saccades. Saccades
direct the fovea, the region of the eye with the highest spatial
imaging resolution, towards putative targets of interest. Mammals make
frequent saccades; many thousands every day. These movements are often
described as ballistic: They are commanded by activity in the superior
colliculus which encodes an endpoint for the saccade trajectory
(\cite{wurtz_superior_1971,hepp_monkey_1993}); once initiated, the
saccade proceeds in a relatively direct and stereotyped (though not
always straight) trajectory (\cite{van_der_stigchel_eye_2006}) to
the endpoint.
% Why am I mentioning curved saccades? Because it shows that saccades
% can exhibit rich behaviour and that modelling them is worthwhile.
However, numerous studies have shown that saccade trajectories are
affected by distractors (\cite{mcpeek_competition_2003,
port_sequential_2003, van_der_stigchel_recent_2010}) and may be
modified in-flight to the extent that saccades may be seen to perform
significant, sometimes even U-shaped, curves towards or away from
distractors. This shows that whilst saccades are simple compared to
the movements of limbs, they still provide information about a number
of mechanisms within the brain and an effort to model them accurately
is worthwhile.

% Say what we've done here, in short form.
Parkinson's Disease is a progressive neurological disorder whose main
symptoms relate to motor control. These include resting tremor,
slowness of movements (bradykinesia), difficulty in initiating
movement (akinesia) and rigidity (refs). It is considered to be a
disease of the basal ganglia because during its progression, dopamine
producing neurons in the substantia nigra pars compacta die away
(ref), affecting the the striatum which depends on endogenous dopamine
for correct operation (ref). The basal ganglia plays an important role
in decision making and action selection (ref), including the selection
of saccadic eye movements (\cite{hikosaka_role_2000}). In the later
stages of Parkinson's Disease, patients show more latent saccades to
targets of interest (ref). They have also been documented as showing
poorer performance in the so-called anti-saccade protocol, in which
the subject is instructed to make a saccade \e{away} from a visual
target. Parkinson's Disease patients make significantly more errors;
they will saccade \e{to} the visual target rather than \e{away} from
it more often than healthy control subjects. To investigate the way in
which degradation in the basal ganglia may be causing these defects we
have developed an integrated brain and biomechanical eye model which
we describe here. The model will initially be used to investigate the
deterioration in anti-saccadic movements which is detectable even in
relatively early-stage Parkinson's Disease.

%See \cite{Rottach_dynamic_1996} for trajectory information in saccades.

@@

Many neural populations are involved in the coding of saccadic eye
movements, only a very brief overview is given here; for a recent
review, see \cite{munoz_commentary:_2002}. One pathway takes
information from the retina directly into the superficial layers of
the superior colliculus in the brainstem (\cite{sterling_receptive_1971,linden_massive_1983,wu_involvement_1994}).
Activity within
the superior colliculus then excites neurons in the pons, medulla and
rostral mid-brain (\cite{sparks_brainstem_2002}) and then through
channels to the motor neurons which innervate the extraocular muscles
(\cite{fuchs_firing_1970,sparks_brainstem_2002}). This direct pathway
is responsible for the low latency saccades called express saccades
(\cite{schiller_effect_1987,edelman_activity_1996}).

% FEF: \cite{schall_neural_1999}
Information from the retina is also processed by visual cortex which
feeds through to the frontal eye fields in which activity is related
to reflexive and voluntary
saccades \cite{schall_neural_1999}. Activity build up in the frontal
eye fields is transferred to the intermediate layers of the superior
colliculus
\cite{stanton_frontal_1988-1} and is also processed by the basal ganglia, which
participates in the selection of the winning saccade end point
\cite{stanton_frontal_1988}.

% On fact that SC doesn't encode oblique/torsional movements:
% \cite{hepp_monkey_1993}

% That means that you need both the expression of movement, and also a
% means by which you can progress the disease in the model.

%If such comparisons are to be made, then the behavioural output of the
%model should be in the form of movements which can be directly
%compared with the biological system. In order to express movements,
%the brain model needs to drive biomechanical neuromuscular models.

%\begin{methods}
\section{Material \& Methods}

% A very short introduction to the section.
The integrated brain and biomechanical model described here is a
development of the model in \cite{cope_basal_2017},
referred to here as the \ccg~model. This was a rate-coded neural
network model incorporating retinal populations, frontal eye fields
(FEF), the basal ganglia (BG), and the superior colliculus (SC). In
the \ccg~model, the centroid of the activity in the deep
layers of superior colliculus was assumed to accurately encode the
location of the eye at the end of the
saccade~(\cite{wurtz_activity_1972,robinson_eye_1972,van_gisbergen_collicular_1987,mcilwain_lateral_1982}).
This location was used to recalculate the positions of the luminances in
the eye's frame of reference at each time step. The model included no
brainstem populations other than superior colliculus, nor a
neuromuscular model.

@@

To the \ccg~model, we have added a rate-coded brainstem model (a `saccadic
burst generator'), and a biomechanical eye, implemented using the
biomechanical modelling framework OpenSim. These will be described below,
but first we will give a description of the co-ordinate systems and the
modelling framework used.

@@

Before describing the biomechanical eye and the brain model, which consisted
of retinotopically mapped neural sheets, we describe the co-ordinate system
used in the world. The eye was located at the origin of a 3 dimensional, right-handed
Cartesian co-ordinate system, with its fovea directed in the $-z$ direction.
There was a notional spherical screen which was also centered at the origin of the
co-ordinate system and had a radius of 50 (in arbitrary units). The \e{fixation
point} was the point on the
screen at which the eye was initially directed.
Onto the screen were projected luminances, each of which having a position
described by two co-ordinates; $\theta_{x}^{t}$, a
rotation of the horizon plane about the $x$ axis, and $\theta_{y}^{t}$, a rotation
of the meridian plane about the $y$ axis. The position is the intersection
of these rotated planes with the spherical screen (disregarding
the intersection point of these three surfaces behind the eye).
%
Note that a luminance with positive $\theta_{x}^{t}$ was above the horizon of this world;
one whose $\theta_{y}^{t}$ was positive lay to the left of the world's meridian. For this
reason, many of the figures in this paper are plotted with $-\theta_{y}$ on the
$x$-axis and $\theta_{x}$ on the $y$-axis so that targets that lay up and to the right
in the world do so in the graphs, also.

Luminances were crosses of height and width subtending $\pm$3\dg~and whose
`bars' were 2\dg~thick. Luminances were oriented like $+$ symbols with
their vertical bar aligned with the meridian plane and their horizontal bar
aligned with the horizon plane described above.

@@

% Say the \ccg~model was implemented in matlab-brahms & this one is SpineML.
The \ccg~model was originally developed to run on the BRAHMS model
execution framework
(\cite{mitchinson_brahms:_2010,mitchinson_brahms_2015}). To run a
BRAHMS model, the researcher must develop \e{BRAHMS components} for
the various neural elements. A BRAHMS component is a programmatically
coded implementation of the behaviour of the component. It may have an
arbitrary number of inputs and outputs and may be written in C, C++,
Python or MATLAB. The \ccg~model's components were hand written in C++
and MATLAB. A BRAHMS \e{SystemML} file describes how the different
components connect together and how data is passed between them
(\cite{mitchinson_brahms:_2010}). The main BRAHMS program first
reads the SystemML file, then dynamically loads all the required
components before executing the system.

In the current work, the \ccg~model was reproduced using the
declarative SpineML markup language (\cite{richmond_model_2014}),
with the help of the graphical SpineML model editing software
called SpineCreator
(\cite{cope_spinecreator:_2016}). SpineML,
which is a development of the NineML specification
(\cite{incf_task_force_on_multi-scale_modeling_network_2011}),
describes neural populations and their projections in a highly
structured format in which neuron bodies, pre- and post-synapses are
described in terms of \e{SpineML components}. These are similar to the
components provided by BRAHMS, but in this case, the components are an
XML description of the functionality of the component, rather than a
programmatic implementation, with one XML file per component. A
SpineML \e{network layer} file then describes which components are
used in the model, and how they are connected together. Finally, a
number of SpineML \e{experiment layer} files specify how the model
described in the network layer can be executed. In the experiment
layer, the execution duration and timestep can be specified, along
with input conditions, connection lesions and component parameter
updates. A complete description of SpineML is given
in \cite{richmond_model_2014}.

SpineCreator, in its r\^ole as a graphical editor for the SpineML
format, was used to generate the SpineML files describing the
model. It was also used to generate the diagrams of the model
(e.g. Fig.~\ref{fig:brain_model}).

As a declarative format for model specification, SpineML is agnostic
about how the model is executed. A number of simulation
engines can be utilised, including DAMSON (\cite{richmond_damson_2015}), GeNN
(\cite{nowotny_flexible_2011,nowotny_spineml_2014}) and BRAHMS (used
here). The simulation engine incorporating BRAHMS is called \stob~
(\cite{cope_spineml_2_brahms_2015-1}). \stob~is a collection of XSLT
stylesheets which first generate and compile C++ BRAHMS components
from the SpineML component layer description files. \stob~then uses
the SpineML network and experiment layer files to generate a BRAHMS
SystemML description of the model. Finally, \stob~executes the model,
now described entirely as a BRAHMS system, via a call to the BRAHMS
binary. A number of additional hand-written components are present
in \stob~providing the inputs (constant inputs, time-varying inputs,
etc) which the modeller specifies in the experiment layer.

In addition to the brain model components, all of which are
code-generated using \stob~as described above, two hand-written
components are integrated into the model: The biomechanical eye model
and a sensory input component. The sensory input component takes the
eye's rotational state and the state of the experimental luminances
and projects a retinotopic activity map into the brain model. Both of
these BRAHMS components were hand-written in C++. To incorporate these
components into the SpineML model, a model-specific modification was
applied in \stob~(\cite{cope_spineml_2_brahms_2015}) making use of
a \stob~feature designed for this purpose (The external.xsl file).

Figure \ref{fig:model_framework} shows the workflow, in which the
model specification files (blue box - a combination of SpineML files
and C++ code), are processed (green box) into a BRAHMS system (red
box).

@@

The brain model, excluding the brainstem, is a re-implementation of
the \ccg~model. Figure \ref{fig:brain_model} shows the layout of the
populations and the interconnections between them. Each population of
2500 elements is arranged as a 50 by 50 grid which forms a map in a
retinotopic co-ordinate system, roughly matching the known layout of
the superior colliculus~(\cite{robinson_eye_1972}). % FIXME: and frontal eye field?

@@

The output, $y$, of LINlinear is related to the activation $a$ by the
piecewise linear transfer function
\begin{equation}
   y(a) = \begin{cases}
      0   & a < c \\
      a-c & c \leq a \leq 1+c \\
      1   & a > 1+c
   \end{cases}
\end{equation}
where $c$ is a parameter defining the slope of the transfer
function. At this point, the naming scheme for the component becomes
apparent; this is a Leaky Integrator with a linear transfer function.

The LINret component used for the retinal populations is again similar
to the LINlinear component, but with no intrinsic noise and no
shunting inhibitory input. It has a neural input which is identical to
the activation input $A$:
\begin{equation}
   a_{in} = A
\end{equation}
The LINexp component is a leaky integrator with an exponential
transfer function. It shares the same differential equation with
LINlinear, but has a different input equation and a different output
transfer function. It has the following equation for the neural
element input $a_{in}$:
\begin{equation}
   a_{in} = [A+N(a-V_{r}^{-})] (1-S) + 0.01 R_N
\end{equation}
where $A$ is the activation input and $N$ is an input which is
modulated by $V_{r}^{-}$, a reversal potential, and $a$, the current
activation of the element. These inputs are summed and then reduced by
a factor which is dependent on $S$, the shunting input. As in
LINlinear, $R_N$ introduces normally distributed noise to the element.

The output, $y$, of the LINexp component is given by
\begin{equation}
   y(a) = \begin{cases}
      \mathrm{e}^{a}-0.9   & \mathrm{e}^a \leq 1+0.9 \\
      1   & \mathrm{e}^a > 1+0.9
   \end{cases}
\end{equation}
This component is used in the STN population, as it gives a more
physiologically accurate f-I
behaviour \cite{wilson_model_2004,bevan_mechanisms_1999,hallworth_apamin-sensitive_2003}
which has been shown to allow the mapping of the basal ganglia network
architecture onto an optimal decision making
model \cite{bogacz_basal_2007}.

>>>>

\section{}
To date, realistic models of how the central nervous system governs
behaviour have been restricted in scope to the brain, brainstem or
spinal column, as if these existed as disembodied organs.
Further, the model is often exercised in relation to an in vivo
physiological experiment with input comprising an impulse, a periodic
signal or constant activation, and output as a pattern of neural
activity in one or more neural populations. Any link to behaviour is
inferred only indirectly via these activity patterns.
%
We argue that to discover the principles of operation of neural
systems, it is necessary to express their behaviour in terms of
physical movements of a realistic motor system, and to supply inputs
that mimic sensory experience. To do this with confidence, we must
connect our brain models to neuro-muscular models and provide relevant
visual and proprioceptive feedback signals, thereby closing the loop
of the simulation.
%
This paper describes an effort to develop just such an integrated
brain and biomechanical system using a number of pre-existing
models. It describes a model of the saccadic oculomotor system
incorporating a neuromuscular model of the eye and its six extraocular
muscles. The position of the eye determines how illumination of a
retinotopic input population projects information about the location
of a saccade target into the system.
%
A pre-existing saccadic burst generator model was incorporated into
the system, which generated motoneuron activity patterns suitable for
driving the biomechanical eye. The model was demonstrated to make
accurate saccades to a target luminance under a set of environmental
constraints.
%
Challenges encountered in the development of this model showed the
importance of this integrated modelling approach. Thus, we exposed
shortcomings in individual model components which were only apparent
when these were supplied with the more plausible inputs available in a
closed loop design. Consequently we were able to suggest missing
functionality which the system would require to reproduce more
realistic behaviour.
%
The construction of such closed-loop animal models constitutes a new
paradigm of \e{computational neurobehaviour} and promises a more
thoroughgoing approach to our understanding of the brain's function as
a controller for movement and behaviour.

@@

% New introduction, tailored to the paradigm shift title.

% Background
The field of computational neuroscience has provided
many \e{systems models} of the brain \citep{arai_two-dimensional_1994,gancarz_neural_1998,hazy_towards_2007,blenkinsop_frequency_2017}.
We refer to these as \e{mechanistic computational models}, meaning
models which consist of populations of neural elements, interconnected
in a biologically plausible manner, which simulate the operation of
the brain. Whilst they differ in scale and complexity, these models
all seek to
describe the fundamental mechanisms behind common animal behaviours
such as locomotion, threat evasion, reaching or feeding. However,
none of the models cited here actually reproduce these
behaviours. In each case, the activity in a certain population of
neurons is taken to be representative of a behavioural outcome.
% Sometimes, ok to just have a brain model
In some cases, it \e{is} reasonable to take the activity of an internal
population within the brain model as being representative of the
induced behaviour. For example, a choice made in a \e{go/no-go} task
could be determined from activity in a population within a basal
ganglia model
\citep{nambu_discharge_1990,kuhn_event-related_2004}. The decision
to \e{go} is selected by a reduction of activity in this population;
maintenance of activity implies \e{no-go}. To validate the model, the
error rates which it generates
could be compared with experimentally determined error rates in
primate subjects. We refer
to this as an \e{output assumption model} because the output is assumed
to signify behaviour. (An \e{input assumption model} assumes that sensory
input produces some particular form of neural activity in an input
population of the model.)

% Gap
However,
we may be interested in reproducing accurate simulated \e{trajectories},
in order to find out how degradation of parts of the model affect
movement. In Parkinson's Disease, degradation of the
dopamine neurons originating in the substantia nigra pars compacta (SNc)
causes diskinesia \citep{galvan_pathophysiology_2008}, as well as
abnormal network activity in
the basal ganglia \citep{brown_dopamine_2001,mccarthy_striatal_2011}.
Sufferers of the disease would be
expected to produce abnormal decision-making \e{and} movement trajectories
in a reach-to-the-correct-target task such as the one described in
\cite{james_target-distractor_2017}.
%
% Solution
A model which
sought to explore in detail the effects of the SNc degradation both on
the decision making \e{and} on the movement dynamics would need a
physically accurate virtual arm, as well as physically realistic sensory
input for the brain. This is no less than a complete model of those sections
of the brain and body which act to fulfil the task. Such a modelling
effort, if successful, would result in a virtual robot capable of expressing
behaviour \e{in response to sensory input from its environment}. This would
represent a paradigm shift in the field of computational neuroscience
worthy of the new name of \e{computational neurobehaviour}.

% Existing models don't control movement unless they control either a mechanical
% or virtual body.
% Can't observe how learning in the model may effect it's evolution in time
% We may miss features about the system
% Brains exist to move muscles

% Solution, build an integrated model. Also give background on OMM system here.

@@

% Solution described means we now need to do background of eye movements:
In an attempt to build a model combining brain, realistic
biomechanics \e{and} sensory
feedback, we sought to extend our previous work modelling the oculomotor system.
The existing model \citep{cope_basal_2017} is already able to capture sensory input
and convert it into a neural signal, assumed to specify the target of a
\e{saccadic eye movement};
a fast movement of the eyes which directs the fovea to a region of interest
in the field of view.
%
The oculomotor system is an excellent candidate for modelling because
its movements can  be specified with only three degrees of freedom,
making it one of the simplest neuro-muscular system in the body.
%
It is nevertheless behaviourally
interesting, as saccadic eye movements reveal information about decision making at
a subconscious level
\citep{deubel_saccade_1996,reppert_modulation_2015,marcos_determining_2016}.
The modelling of the oculomotor system is served by a large body of
behavioural data describing
saccades \citep{tipper_reaching_2001,walker_effect_1997,casteau_effect_2012},
many anatomical studies of the neural substrates involved
\citep{meredith_intrinsic_1998,isa_intrinsic_2002,isa_exploring_2009}
and electrophysiological data linking these together
\citep{hepp_spatio-temporal_1983,dorris_neuronal_1997,mcpeek_competition_2003,vokoun_circuit_2011}.
Furthermore, in the context
of building \e{behaving} systems a
necessary part of any model for which the behaviour requires visual attention
and decision making is a realistic mechanism for gathering visual information.
This is obvious from extrinsic considerations---a
subject must look at a scene to make decisions or navigate within it. It
also follows for \e{intrinsic} reasons. For example, \cite{howard_hand_1997}
showed that visual cues affect reach trajectories and the same group later
demonstrated that reaching affects the saccadic system
\citep{tipper_reaching_2001} suggesting a close relationship between these
neural systems. Building the simplest, behaving oculomotor system will
therefore assist future computational neurobehavioural modelling efforts.

Many neural populations are involved in the coding of saccadic eye
movements, only a very brief overview is given here; for a
review, see \cite{munoz_commentary:_2002}. One pathway takes
information from the retina directly into the superficial layers of
the superior colliculus in the brainstem
\citep{sterling_receptive_1971,linden_massive_1983,wu_involvement_1994}.
Activity within
the superior colliculus then excites neurons in the pons, medulla and
rostral mid-brain \citep{sparks_brainstem_2002} and then through
channels to the motor neurons which innervate the extraocular muscles
\citep{fuchs_firing_1970,sparks_brainstem_2002}. This direct pathway
is responsible for the low latency saccades called express saccades
\citep{schiller_effect_1987,edelman_activity_1996}.
%
Information from the retina is also processed by visual cortex which
feeds through to the frontal eye fields in which activity is related
to reflexive and voluntary
saccades \citep{schall_neural_1999}. Activity build-up in the frontal
eye fields is transferred to the intermediate layers of the superior
colliculus
\citep{stanton_frontal_1988-1} and is also processed by the basal ganglia,
which participates in the selection of the winning saccade end point
\citep{stanton_frontal_1988}.
%
Although both cortical and subcortical paths produce a saccade
target signal in the superior colliculus, it is also possible for
animals to make relatively normal saccades even after the
colliculus has been ablated
\citep{wurtz_activity_1972,aizawa_reversible_1998},
though express saccades are lost with collicular lesions
\citep{schiller_effect_1987}. This makes the superior colliculus
a perplexing structure, being both critically involved in saccade
target specification \citep{sparks_sensory_1987} and saccade dynamic control
\citep{waitzman_superior_1991,goossens_optimal_2012} and yet dispensible.
The `backup pathway' likely incorporates the oculomotor vermis and
fastigial oculomotor region of the
cerebellum which are known to participate in the specification, dynamics
and adaptation of saccadic eye movements
\citep{kleine_saccade-related_2003,takagi_effects_1998}.

There is a long history of modelling the oculomotor system. For a comprehensive
review, see \cite{girard_brainstem_2005}. Models of individual systems have been proposed for brainstem \citep{robinson_oculomotor_1975,scudder_new_1988,gancarz_neural_1998},
cerebellum \citep{quaia_model_1999,dean_modelling_1995,dean_learning_1994} and
superior colliculus
\citep{arai_two-dimensional_1994,moren_mechanism_2013,marino_spatial_2012}.
More recently, combined models have also been
developed incorporating sensory input \citep{cope_basal_2017} and driving simple
physical plants representing the eye \citep{tabareau_geometry_2007,nguyen_saccade_2014,thurat_biomimetic_2015}.
None of these models has yet fully closed the loop to produce a behaving system
operating freely within its environment. We argue that developing
integrated, closed-loop models of behaving systems offers insights into
the operation of neural systems that are not available from
input-~or output-assumption models.

\section{Material \& Methods}

% A very short introduction to the section.
The integrated brain and biomechanical model described here is a
development of the model in \cite{cope_basal_2017},
referred to here as the \ccg~model. This was a rate-coded neural
network model incorporating retinal populations, frontal eye fields
(FEF), the basal ganglia (BG), and the superior colliculus (SC). In
the \ccg~model, the centroid of the activity in the deep
layers of superior colliculus was assumed to accurately encode the
location of the eye at the end of the
saccade \citep{wurtz_activity_1972,robinson_eye_1972,van_gisbergen_collicular_1987,mcilwain_lateral_1982}.
This location was used to recalculate the positions of the luminances in
the eye's frame of reference at each time step. The model included no
brainstem populations other than superior colliculus, nor a
neuromuscular model.

@@

To the \ccg~model, we added a rate-coded brainstem model (a `saccadic
burst generator') and a biomechanical eye, implemented using the
biomechanical modelling framework OpenSim \citep{seth_opensim_2015}.
These will be described below, but first we will give a description of
the co-ordinate systems and the modelling framework used.

@@

Before describing the biomechanical eye and the brain model, which consisted
of retinotopically mapped neural sheets, we describe the co-ordinate system
used in the world. The eye was located at the origin of a three-dimensional, right-handed
Cartesian co-ordinate system, with its fovea directed in the $-z$ direction.
There was a notional spherical screen which was also centred at the origin of the
co-ordinate system and had a radius of 50 (in arbitrary units). The \e{fixation
point} was the point on the
screen at which the eye was initially directed.
Onto the screen were projected target luminances, each of which having a position
described by two co-ordinates; $\theta_{x}^{t}$, a
rotation of the horizon plane about the $x$ axis, and $\theta_{y}^{t}$, a rotation
of the meridian plane about the $y$ axis. The position is the intersection
of these rotated planes with the spherical screen (disregarding
the intersection point of these three surfaces behind the eye).
%
Note that a luminance with positive $\theta_{x}^{t}$ was above the horizon of this world;
one whose $\theta_{y}^{t}$ was positive lay to the left of the world's meridian. For this
reason, many of the figures in this paper are plotted with $-\theta_{y}$ on the
$x$-axis and $\theta_{x}$ on the $y$-axis so that targets that lay up and to the right
in the world do so in the graphs, also.

Luminances were crosses of height and width subtending $\pm$3\dg~and whose
`bars' were 2\dg~thick. Luminances were oriented like $+$ symbols with
their vertical bar aligned with the meridian plane and their horizontal bar
aligned with the horizon.

@@

% Say the \ccg~model was implemented in matlab-brahms & this one is SpineML.
The \ccg~model was originally developed to run on the BRAHMS model
execution framework
\citep{mitchinson_brahms:_2010,mitchinson_brahms_2015}. To run a
BRAHMS model, the researcher must develop \e{BRAHMS components} for
the various neural elements. A BRAHMS component is a programmatically
coded implementation of the behaviour of the component. It may have an
arbitrary number of inputs and outputs and may be written in C, C++,
Python or MATLAB. The \ccg~model's components were hand written in C++
and MATLAB. A BRAHMS \e{SystemML} file describes how the different
components connect together and how data is passed between them
\citep{mitchinson_brahms:_2010}. The main BRAHMS program first
reads the SystemML file, then dynamically loads all the required
components before executing the system.

In the current work, the \ccg~model was reproduced using the
declarative SpineML markup language \citep{alex_cope_spineml_2014,richmond_model_2014},
with the help of the graphical SpineML model editing software
called SpineCreator
\citep{cope_spinecreator_2015,cope_spinecreator:_2016}. SpineML,
which is a development of the NineML specification
\citep{incf_task_force_on_multi-scale_modeling_network_2011},
describes neural populations and their projections in a highly
structured format in which neuron bodies, pre- and post-synapses are
described in terms of \e{SpineML components}. These are similar to the
components provided by BRAHMS, but in this case, the components are an
XML description of the functionality of the component, rather than a
programmatic implementation, with one XML file per component. A
SpineML \e{network layer} file then describes which components are
used in the model, and how they are connected together. Finally, a
number of SpineML \e{experiment layer} files specify how the model
described in the network layer can be executed. In the experiment
layer, the execution duration and timestep can be specified, along
with input conditions, connection lesions and component parameter
updates. A description of SpineML is given
in \cite{richmond_model_2014}; the definitive definition is found
in the schemas \citep{cope_spineml_2014}.
SpineCreator, in its r\^ole as a graphical editor for the SpineML
format, was used to generate the SpineML files describing the
model. It was also used to generate the diagrams of the model.

As a declarative format for model specification, SpineML is agnostic
about how the model is executed. A number of simulation
engines can be utilised, including DAMSON \citep{richmond_damson_2015}, GeNN
\citep{nowotny_flexible_2011,nowotny_spineml_2014} and BRAHMS (used
here). The simulation engine incorporating BRAHMS is called \stob~
\citep{cope_spineml_2_brahms_2015-1}. \stob~is a collection of XSLT
stylesheets which first generate and compile C++ BRAHMS components
from the SpineML component layer description files. \stob~then uses
the SpineML network and experiment layer files to generate a BRAHMS
SystemML description of the model. Finally, \stob~executes the model,
now described entirely as a BRAHMS system, via a call to the BRAHMS
binary. A number of additional hand-written components are present
in \stob~providing the inputs (constant inputs, time-varying inputs,
etc) which the modeller specifies in the experiment layer.

In addition to the brain model components, all of which are
code-generated using \stob~as described above, two hand-written
components are integrated into the model: The biomechanical eye model
and a sensory input component. The sensory input component takes the
eye's rotational state and the state of the experimental luminances
and projects a retinotopic activity map into the brain model. Both of
these BRAHMS components were hand-written in C++. To incorporate these
components into the SpineML model, a \stob~\e{external.xsl} file was used.
The external.xsl file scheme for incorporating external BRAHMS components
into a SpineML model was a new \stob~feature motivated by the current work.
Fig.~\ref{model_framework} shows the workflow, in which the
model specification files (blue box - a combination of SpineML files
and C++ code), are processed (green box) into a BRAHMS system (red
box).

@@

The brain model, excluding the brainstem, is a re-implementation of
the \ccg~model. Fig.~\ref{brain_model} shows the layout of the
populations and the interconnections between them. Each population of
2500 elements is arranged as a 50 by 50 grid which forms a map in a
retinotopic co-ordinate system, roughly matching the known layout of
the superior colliculus \citep{robinson_eye_1972}. % FIXME: and frontal eye field?

@@

The output, $y$, of LINlinear is related to the activation $a$ by the
piecewise linear transfer function
\begin{equation}
   y(a) = \begin{cases}
      0   & a < c \\
      a-c & c \leq a \leq 1+c \\
      1   & a > 1+c
   \end{cases}
\end{equation}
where $c$ is a parameter defining the slope of the transfer
function. At this point, the naming scheme for the component becomes
apparent; this is a Leaky Integrator with a piecewise-linear transfer function.

The LINret component used for the retinal populations is similar
to the LINlinear component, but with no intrinsic noise and no
shunting inhibitory input. It has a neural input which is identical to
the activation input $A$:
\begin{equation}
   a_{in} = A
\end{equation}
The LINexp component is a leaky integrator with an exponential
transfer function. It shares the same differential equation with
LINlinear, but has a different input equation and a different output
transfer function. It has the following equation for the neural
element input $a_{in}$:
\begin{equation}
   a_{in} = [A+N(a-V_{r}^{-})] (1-S) + 0.01 R_N
\end{equation}
where $A$ is the activation input and $N$ is an input which is
modulated by $V_{r}^{-}$, a reversal potential, and $a$, the current
activation of the element. These inputs are summed and then reduced by
a factor which is dependent on $S$, the shunting input. As in
LINlinear, $R_N$ introduces normally distributed noise to the element.

The output, $y$, of the LINexp component is given by
\begin{equation}
   y(a) = \begin{cases}
      \mathrm{e}^{a}-0.9   & \mathrm{e}^a \leq 1+0.9 \\
      1   & \mathrm{e}^a > 1+0.9
   \end{cases}
\end{equation}
This component is used in the subthalamic nucleus (STN) population, as it gives a more
physiologically accurate f-I
behaviour \citep{wilson_model_2004,bevan_mechanisms_1999,hallworth_apamin-sensitive_2003}
which has been shown to allow the mapping of the basal ganglia network
architecture onto an optimal decision making
model \citep{bogacz_basal_2007}.

The D1MSN and D2MSN components are both leaky integrators, similar to
LINlinear. They differ in that they have no shunting inhibition. They
are used in to model medium spiny neuron (MSN) populations in
the striatum. As they model
the fact that most MSN neurons fall into two groups; those expressing D1
dopamine receptors and those expressing D2 receptors, they have a
dopamine parameter that modulates the input activation, so
that their equations for $a_{in}$ are thus:
\begin{equation}
   a_{in}^{D1} = (0.2 + d)A + 0.01 R_N
\end{equation}
\begin{equation}
   a_{in}^{D2} = (1 - d)A + 0.01 R_N
\end{equation}
where $d$ is the dopamine parameter. Varying dopamine from 0 to 1
enhances the activation in the D1 model, whereas it decreases
the activation of the D2 model elements, in line with experimental
observations \citep{harsing_influence_1997,gonon_prolonged_1997}.
Note that the equation for $a_{in}^{D1}$
differs from that used in the \ccg~model, for which the
cortico-striatal weights are multiplied by $(1+d)$ rather than
$(0.2+d)$.
